Testing Charts: viewer's perceptual accuracy in surveys

### 2023-07-17 SDSS-2023/testing-charts-accuracy.qmd
move html to default position

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: 1               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Heike
    first-init: H.
    last: Hofmann
    marker: 2
  - first: Nola
    first-init: N.
    last: duToit
    marker: 1
  - first: Edward
    first-init: E.
    last: Mulrow
    marker: 1
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: National Opinion Research Center (NORC)
    country: United States
    marker: 1
  - institution: Iowa State University
    department: Statistics
    country: United States
    marker: 2
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
  \usepackage{booktabs}
  \usepackage{graphicx}
format:
  html: default
editor: 
  markdown: 
    wrap: sentence
---

```{=latex}
  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}
  \setlength{\parindent}{0pt}
```
\hh{hello?}

\kr{Kiegan, this is your editing color :) - you can choose a different color above}

<!-- fig-width: 6.5 -->

<!-- fig-height: 4.5 -->

<!-- template: template-jdsart.cls.tex -->

@@

Participants were recruited as part of NORC's AmeriSpeak panel, which utilizes a probability-based sampling methodology and samples U.S. households from NORC's National Sample Frame that provides coverage of over 97% of U.S. households.
The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
Each test was conducted using the AmeriSpeak OMnibus survey, which runs biweekly and pulls a sample of around 1,000 U.S. adults and asks participants to answer questions on a variety of topics.
The advantage of using a probability-based approach is two-fold.
First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.

@@

```{r}
#| label: fig-tasks-12
#| fig-cap: "The two stacked barcharts every participant got to see. In each barchart, the two marked tiles are to be compared for their size. In both instances, the tile on the right is (very slightly) larger."
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| out-width: 100%
knitr::include_graphics("images/vertical-ab.PNG")
knitr::include_graphics("images/vertical-cd.PNG")
```

@@

-   When combining responses from multiple rounds, we make weighting adjustments to ensure that each weights in each sample are properly calibrated to the population total across both rounds.
    **NEED ED REVIEW OF THIS LANGUAGE.** This adjustment is completed as follows: \[insert Heike text from analysis doc here\].

@@

-   Design choices impact viewer behavior

-   Design choices impact viewer accuracy

-   Reproduced prior convenience sample results using a large, nationally-representative survey population

@@

@@

>>>>

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
short-title: Testing charts in surveys # running title
author:
  - first: Kiegan                 # first name
    first-init: K.              # initial of first name
    last: Rice                  # last name
    marker: 1               # identifier for the affiliation
    footnote-id: 1              # optional identifier for the footnote
    email: foo@bar.com          # optional email address
  - first: Heike
    first-init: H.
    last: Hofmann
    marker: 2
  - first: Nola
    first-init: N.
    last: duToit
    marker: 1
  - first: Edward
    first-init: E.
    last: Mulrow
    marker: 1
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: National Opinion Research Center (NORC)
    country: United States
    marker: 1
  - institution: Iowa State University
    department: Statistics
    country: United States
    marker: 2
bibliography: references.bib
abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work."
keywords: 'list, keywords, here, 3-6, are, fine'
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
  \usepackage{booktabs}
  \usepackage{graphicx}
  \usepackage{hyperref}
format:
  html: default
  pdf: 
    documentclass: jds
    include-in-header: authors.tex
    keep-tex: true
editor: 
  markdown: 
    wrap: sentence
---

```{=latex}
  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}
  \setlength{\parindent}{0pt}
```

<!-- fig-width: 6.5 -->
<!-- fig-height: 4.5 -->
<!-- template: template-jdsart.cls.tex -->

@@

Participants were recruited as part of NORC's AmeriSpeak panel, which utilizes a probability-based sampling methodology and samples U.S. households from NORC's National Sample Frame that provides coverage of over 97% of U.S. households.
The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.adults to answer questions on a variety of topics.
The advantage of using a probability-based approach is two-fold.
First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.

\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?}

@@

```{r}
#| label: fig-tasks-12
#| fig-cap: "Two stacked (vertical) barcharts. In each barchart, two tiles are marked. In both instances, the tile on the right is (very slightly) larger. "
#| fig-subcap:
#| - "**Unaligned bars**"
#| - "**Aligned bars**"
#| layout-nrow: 1
#| out-width: 100%
knitr::include_graphics("images/vertical-ab.PNG")
knitr::include_graphics("images/vertical-cd.PNG")
```

```{r}
#| label: fig-tasks-3
#| fig-cap: "Changes to to the vertical design are made in two steps. First, the original chart is rotated. The areas of all tiles are kept the same. In a second step, the aspect ratio of tiles is changed while keeping the areas of tiles the same."
#| fig-subcap:
#| - "**Tall horizontal bars**, unaligned"
#| - "**Wide horizontal bars**, unaligned"
#| layout-nrow: 1
#| out-width: 100%
knitr::include_graphics("images/round3/horizontal-ab.PNG")
knitr::include_graphics("images/round3/hwide-ab.PNG")
```

@@

-   When combining responses from multiple rounds, we make weighting adjustments to ensure that weights in each sample are properly calibrated,  to the population total across both rounds.
    **NEED ED REVIEW OF THIS LANGUAGE.** This adjustment is completed as follows: \[insert Heike text from analysis doc here\].

@@

@@

-   Reproduced (some) prior convenience sample results using a large, nationally-representative survey population:

    - Cleveland: position (aligned) versus unaligned
    - unaligned tiles within stacked bar are a hybrid between framed bar and floats
    - Talbot: pie charts - individual wedges vs piecharts

-   Design choices impact viewer accuracy:

    - new finding: pie charts accuracy in work better when framed (wedges)

- Other important measures beyond accuracy: time to completion (automatic on web), certainty in response (asked).

@@

    - Design choices impact viewer behavior: zooming

- (some) Demographics matter: age and gender does not seem to matter for perception
  economic background and education does matter for perception: we need to know, we do not want to create a hurdle in communication
